Internal Gravity Wave Breaking in White Dwarf Binaries

### 2019-04-29 writeups/paper.tex

\begin{abstract}
    In white dwarf binaries tidal torques are expected to raise a train of
    internal gravity waves in the white dwarf interior. These
    outward-propagating waves can then grow to be nonlinear and break via
    hydrodynamical instabilities, transfering energy and angular momentum from
    the binary orbit to the white dwarf. We perform 2D numerical simulations of
    nonlinear wave breaking of outwards-propagating internal gravity waves in an
    incompressible, isothermal atmosphere with an exponentially decaying density
    stratification. We find that tidal synchronization, after an initial
    transient phase, proceeds inwards from the surface in thin layers. We argue
    the thickness of these layers is limited by the Kelvin-Helmholtz
    Instability. We provide simple analytical formulae for the location and
    thickness of tidal synchronization over time that are in good agreement with
    our simulations.
\end{abstract}

>>>>

\begin{abstract}
    In sufficiently compact white dwarf binaries, dynamical tides raise a train
    of internal gravity waves that propagate towards the surface. We perform 2D
    numerical simulations of these waves undergoing nonlinear wave breaking in
    an incompressible, isothermal atmosphere. After an initial transient phase,
    we find that these waves induce a sharp transition between a non-rotating
    core and synchronously rotating envelope. We find evidence that the width of
    this transition layer is bound from below by the Kelvin-Helmholtz
    Instability. We provide analytical formulae for absorption and reflection of
    incident waves off the critical layer up to prefactors of order unity. These
    prefactors converge to constant values when artificial dissipation is
    decreased. We provide dimensionless criteria necessary to resolving momentum
    transfer within the critical layer. Finally, we speculate on the application
    of our model to tidal synchronization and heating in astrophysical systems.
\end{abstract}
